Stage-Splitting a Modal Language

### 2014-10-20 paper/umut.tex
working on the related work

>>>>

\paragraph{Staging and staging transformations.}

To the best of our knowledge, Jorring and Sherlis are the
first to define precisely the notion of {\em staging} by recognizing
it to be a common optimization performed in software design, as the
programmer tries to stage computations based on frequency of a
computation and the availability of data~\cite{jorring86}.  They then
define, staging as the problem of taking a function of the form
$f(x,y)$ and deriving two functions $f_1$ and $f_2$ such that $f(x,y)
= f_2(f_1(x),y)$.  As the authors point out, the key difference
between staging and partial evaluation as that the latter performs no
a priori transformation; instead when $x$ becomes available it
specialized $f(x,\cdot)$ as much as possible.  Based on this notion of
staging, Jorring and Sherlis also define the {\em meta programming}
approach where $f(x, \cdot)$ is specialized by a meta program derived
from the definition of $f(x,y)$.

Having defined staging and meta programming problems, Jorring and
Scherlis move on to describe techniques for them.  They present the
technique called {\em pass separation} as a solution the problem of
splitting a function into two such that the stage-1 function produces
an intermediate data structure from the stage-1 data, which is then
used by the stage-2 function along with stage-2 data to produce the
final result.  Jorring and Scherlis then note describe how
pass-separation can be applied to an interpreter to obtain a compiler
from it by manually transforming the interpreter code into a compiler
by analysis. They also describe how meta programming can be used to
derive a specialized program not by creating a data structure as pass
separation does by generating a more optimized code that ``freezes''
certain actions to be performed at run time.

Jorring and Scherlis highlight that their techniques are not meant to
be applied automatically and that they describe what can be called a
``vision.''  Indeed they describe their approach by way of examples as
transformations that can be applied by the programmer manually and
point out that they are powerful techniques that will require some time
and further research to automate.

Our work provides strong evidence that pass-separation transformations
can indeed be performed on explicitly staged programs automatically.

\ur{I didn't understand meta programming transformations that they do
  and the relationship to followup work.}
